---
title: "Introduction to `smlePH` package"
# output: rmarkdown::html_vignette
author: 
  - Susan Halabi
  - Taehwa Choi
  - Yuan Wu
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Introduction to `smlePH` package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { equationNumbers: { autoNumber: "AMS" } }
});
</script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(smlePH)
```


## Introduction

`smlePH` is the R package to fit the proportional hazards model with a sieve-based maximum full likelihood approach.

Assume that there are $n$ patients in a cohort. Denote $T_i$ as the survival time and $x_i$ as the $p$-dimensional covariates. We observe $(Y_i = \min(T_i, C_i), \delta_i = I(T_i < C_i))$ under the right-censoring, thus the effective sample consists of $\{Y_i = T_i \wedge C_i, \delta_i = I(T_i < C_i), x_i\}_{i=1}^n$.
Let $\beta$ be the $p$-dimensional regression coefficient.
We posit the PH model as follows
\begin{equation}
\label{eq:phmod}
\Lambda(t|x_i) =  \Lambda_0(t) e^{\beta' x_i},~
i=1,\ldots,n.
\end{equation}
where $\Lambda(t|x_i)$ and $\Lambda_0(t)$ are the conditional cumulative hazard function given $x_i$ and the baseline cumulative hazard function, respectively.

Instead of using the partial likelihood approach (Cox, 1972), we utilize a full log-likelihood with respect to $(\beta,\Lambda)$ to estimate \eqref{eq:phmod},
\begin{equation}
\label{eq:fl}
\ell(\beta,\Lambda) = \sum_{i=1}^n \delta_i 
\{ \log \lambda_0(t_i) + \beta'x_i \} - \sum_{i=1}^n
\Lambda_0(t_i) e^{\beta'x_i}.
\end{equation}



## References
